Towards Detecting Stalkerware Through Network Traces

### 2019-03-23 paper/main.tex
Finished run-through number 1

For the apps whose traffic do not contain "spy," "find," or "track," we 
next check for the app name, the manufacturer name, or substrings of either. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption. Manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we can only 
identify by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.

@@

We evaluated both specific and general signatures by taking twenty-four hours 
of network traffic captures on our own personal phones. Our goal was to 
determine a lower bound for our false positive rates. These 
captures also included periods of time that were deemed high risk for false 
positives as they emulated behavior that is most likely to be caught by our 
signatures. Given the relatively small amount of data we have to work with, our 
results should be considered extremely conservative lower bounds on the true 
false positive rates of our signatures. TODO: explain how we got false pos rate

@@

Using the generic filters, we identified two TCP streams and 3 domains out of 
our total of 2869 streams and 1756 domains in our self network capture. In 
addition, these filters will only catch 21 out of the 24 apps that comprise the 
specific filter. This is an issue as it does infer that our generic filters are 
too generic. Upon looking at the specified domains, we see that these are most 
likely adware services that track user browsing data. Although there is a 
possibility that these can be considered malicious enough to add to our 
results, these are not stalkerware and as such count towards our false positive 
rate. Also while the percent of false positives is quite low, if we were to 
deploy this filter in an actual router, it would undoubtedly collect far more 
domains than we should be catching.

While looking at domains and TCP streams gets a good sense of how our filters work in a router, ideally we would also want to see how many apps we misclassify using our generic method. This is because our study is based on stalkerware in phones, and we make no assumption that stalkerware is only on cellular devices. Since we currently cannot connect each domain or TCP stream with a single app without identifying and tracking the traffic for all of our own apps, we make the assumption that each app will connect to only one domain, with the exception of the web browser. Using this assumption, we see that none of the apps on our phones are caught by our filter. Moreover, even apps meant for tracking buses do not get caught by the filter, despite the obvious connection that one could make.

@@

In the future, we would like to automate our process. Ideally we would like to use machine learning to produce signatures for apps we have deemed stalkerware and any of it's copies. Moreover, when we have enough signatures to comprise a training and validation set, we can apply machine learning to fuse these signatures into a compact filter. This would likely be a good amount of work, but if it could be created it would make most stalkerware not in our dataset easily detectable.

Previously, we had expressed an interest in seeing if the signatures we found 
by hand could identify more apps without manual inspection. There are (as far 
as we can see) two possible methods of doing this. The first is to 
auto-generate signatures based on a list of keywords that commonly appear in 
domains, SNIs, User-Agents, and HTTP host names for stalkerware traffic. These 
keywords might include the names of the apps, substrings that appear in the app 
names, and words that seem relevant to tracking apps. A more complicated method 
would be to set up bots that can download potential spyware by searching 
various terms on the app store like "tracker app." The bots would have to run 
the apps and monitor their traffic, identifying app traffic by eliminating 
traffic from other sources. The remaining traffic would belong to the app, and 
the domains, SNIs, user-agents, and any other available identifying features 
could be recorded. However, the most difficult problem in this case becomes 
determining which apps are actually stalkerware on the app store, which is 
beyond the scope of this project.

Additionally, we would like to actually test our signature on real word traffic. We are unable to do so currently since our method currently requires a network capture, and we do not have nor are able to get consent from everyone who uses a given network router. Hopefully we would be able to clear away these considerations in order to find the actual false positive rate of our packet filters.

@@

Stalkerware is a problem, in the context of both society and security. As a tool for partner abuse, its use needs to be prevented before it can cause physical or mental harm, otherwise people cannot trust cell phones for their safety. As a software, current methods are unable to detect stalkerware on any given phone, the method of installation also exposes a security vulnerability that can be exploited for other purposes. In this paper, we have created a method that uses the intrinsic requirement of stalkerware, network communication, to detect it's presence on any device we can link to an IP address. This novel method is able to distinguish stalkerware traffic despite its relative simplicity.

Overall, we find that our method provides a much higher chance of detection than current antivirus software or IDS. We believe that this method can be expanded upon to be either more accurate or more inclusive. This method does have issues that prevent it from actual deployment, however it provides a good step forward in detecting stalkerware.

>>>>

For the apps whose traffic do not contain "spy," "find," or "track," we 
next check for the app name, the manufacturer name, or substrings of either. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption. Manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps whose only 
general signature is their manufacturer are made by GPSWOX and GeoZilla. GPSWOX 
makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.

@@

We evaluated both specific and general signatures by taking 24 hours 
of network traffic captures on our own personal phones. Our goal was to 
determine a lower bound for our false positive rates. These 
captures also included periods of time where we attempted to trigger false 
positives by emulating behavior that might match our 
signatures. For example, we used the UCSD shuttle tracking app in the hopes 
that it would get caught by the keyword method for containing the word 
``track.'' Given the relatively small 
amount of data we have to work with, our 
results should be considered extremely conservative lower bounds on the true 
false positive rates of our signatures. Ideally, we would determine the false 
positive rate by counting how many apps our method misclassifies. However, 
separating traffic by the originating application is a nontrivial problem that 
we were not able to solve in the time available. We settled for two 
approximations: the number of misclassifed TCP streams and the number of 
misclassified domains.

@@

Using the general filters, we identified two TCP streams and three domains out 
of our total of 2,869 streams (0.07\%) and 1,756 domains (0.2\%) in our 24 
hours of personal traffic. In addition, these filters will only catch 21 out of 
our 24 apps, making the general filetrs' false negative rate 12.5\%. The 
misclassified domains are all advertisement services that 
track user browsing data. While the percentage of false positives is quite low, 
we are wary of assigning it too much significance. We predict that if we were 
to deploy this filter in a real router, it 
would undoubtedly collect far more domains than we did with our personal 
traffic.

@@

In the future, we would like to automate our signature identification process. 
If we could acquire enough signatures to comprise a training and validation 
set, we could determine whether machine learning could help us accurately 
distinguish stalkerware from innocent apps. A more complicated method 
would be to set up bots that can download potential spyware by searching 
various terms on the app store like "tracker app." The bots would have to run 
the apps and monitor their traffic, identifying app traffic by eliminating 
traffic from other sources. The remaining traffic would belong to the app, and 
the domains, SNIs, user-agents, and any other available identifying features 
could be recorded. However, the most difficult problem in this case becomes 
determining which apps are actually stalkerware on the app store without 
examining their traffic first.

Additionally, we would like to actually test our signatures on real word 
traffic. We are unable to do so currently since our method currently requires a 
network capture, and we do not have nor are able to get consent from everyone 
who uses a given network router. We would like to find a way to use 
differential privacy (such as RAPPOR \cite{erlingsson_rappor:_2014}) to ensure 
that we can only collect aggregate statistics on 
stalkerware usage, without identifying any device as having it, but have yet to 
find a way to set that up.

@@

Stalkerware is a pervasive and troubling problem, in the context of both 
society and security. As a tool for partner abuse, its use needs to be 
prevented before it can cause physical or mental harm, otherwise people cannot 
trust cell phones for their safety. Current methods of stalkerware detection 
are unreliable, leading to a lack of data on their use. The context in which 
stalkerware is used also invalidates many assumptions about computer security, 
such as the idea that strong passwords and antivirus can keep a phone free of 
malware. We have taken a step towards answering the question of how much 
stalkerware is in existence by creating signatures for network traffic. This 
novel method is able to distinguish stalkerware traffic despite 
its relative simplicity. We also describe several findings about the 
stalkerware ecosystem, the most interesting of which is the knowledge of how 
many apparently unique apps are actually copies of one another. Our methods 
can certainly be expanded upon to be either more accurate or more 
inclusive, but they provide a good step forward in detecting stalkerware.
